
Start with name here


Then maybe table of contents


Introduction

The main goal of this project was to design a protocol to
network a lot of microcontroller using cheap nrf24l01+ modules.  We
decided to develop a self organizing spanning tree mesh network using
a pre-defined node to be sink.  The sink was connected to a pc which
made the mesh network available though a web socket.  The result is
the possibility to make large, cheap, long-range networks which are
accessable though the internet.  Possible use-cases include remote
monitoring and home automation.

Design decisions

The radiomodules support automatic acknoledgement of packets with
optional retransmissions. This enabled us to easily make the
communication between two modules reliable. The auto acknoledgement
feature was used in all communication except for new nodes
broadcasting to connect to the network.

The radio modules uses GFSK in the 2.4-2.5125Ghz range with support of
datarates from 250Kbps up to a maximum of 2Mbps. The modules uses 1Mhz
of bandwidth per channel and support 125 different channels. We
decided to use the same frequency for all nodes in the same network to
decrease complexity. This makes all communication within the network
interfere. The radiomodules has internal addressing with addresses
consisting of 3-5 bytes. Each module can listen for up to six adresses
at the same time but five of these adresses can only differ in their
last byte. We decided to let the mcu save the addresses and routing
table and only use two of the addresses in the radio modules. One for
broadcast messages and one for unicast messages directed towards the
specified node.

We studied different routing protocols for making a mesh-network,
initally we wanted to have a complete mesh network without a sink but
due to the limited amout of memory availible on the chosen mcu we
decided to go for a spanning tree network instead. We figured that
this would make the routing table use less memory since and be less
complex. 

Each node keeps a routing table of all nodes locaded below them in the
network and their corresponding parent node. The sink is the highest
parent. This approach uses memory scaling linearly with the amount of
nodes in the network. The nodes only need to know the route to the
nodes located below themselves.

Hardware: 

Each node consists of one atmega328p mcu running from the
internal oscillator at 8Mhz. It's connected to a nrf24l01+ through the
SPI port. The mcu can be connected to an rs232 adapter to enable
debugging messages to a computer. The sink mcu is connected to an
rs232 adapter to interface with the PC. The nodes accept an input
voltage from 2.4v to 3.6v. Since the NRF24l01+ modules can be run at a
minimum of 1.8v we could potentially run the node at 1.8v but this
would require a lower clockfrequency for the mcu.

Software:

The software on the mcu acts as a big statemachine. At startup the mcu
is in the unconnected stated. In this state the node is completely
unconnected the network. If a node want's to join a network it needs
to send a broadcast which makes it go into the BRD SENT state. As long
as it does not receive any replies the mcu stays in this state and
sends new broadcasts with a fixed interval.

If the node gets a reply from another node it goes into the
collecting parents state. The node stays and listens to the channel for
a fixed amount of time. If more nodes than one replies to the
broadcast the node chooses the one with the smallest hop-count and moves to
either the wait_for_Address state or the connected state.  If this
node has received an address before it will go directly to the connect
state. Otherwise it will ask the network to give it a free address and
go into the wait_for_address state.  ...... continue to describe the
state shiet. Explain why we choose polling or interrupts....


Tests:
We measured the current of a non sink node to 13mA at 3.3v. The
currnet measurement was made from the output of a 3.3v regulator. 
The node was put in a always listening state. According to the 
datasheet the radio module should consume a maximum of 11mA in
listening mode.

We did not to any tests for the power consumption in transmitting
mode.

Maximum range:
We measured the maximum range of one hop to 80m in an outdoor 
enviroment when transmitting at 2.52GHz. 
The maximum indoor range was measured to 70m in an indoor
enviroment @ 2.52GHz.
 
